---
title: "05 Tidy Data and Types Starter Code"
format: docx
editor: visual
---

## Lecture 05 Notes

```{r}

# Load the tidyverse.

```

```{r}

# Read in the customer and transaction data (make sure it's in your working directory).

# join the store transactions data to the customer data - call the joined data "crm_data"

```

### Data Structure and Summarizing

```{r}

# Display each customer's region, ID, and monthly purchase counts for 2018.

crm_data |> 
  select(region, customer_id, contains("2018"))
```

### Pivot Longer

```{r}

# Each row will represent one customer in one month, including zero-purchase months.
# Month names (e.g. "jan_2018") become values in "month_year"; "transactions" holds purchase counts.

crm_long <- crm_data |>
  select(region, customer_id, contains("2018")) |> 
  pivot_longer(
    -c(region, customer_id),
    names_to = "month_year",
    values_to = "transactions"
  )

crm_long
```

```{r}

# Report 2018 purchase activity for each region using crm_long.
# Show the total purchases and average purchases per customer per month,
# including months with zero purchases.
```

### Pivot Wider

```{r}

# This code pivots our long form data frame back to its original wide state.

crm_long |> 
  pivot_wider(
    names_from = month_year,
    values_from = transactions
  )
```

### Your Turn!

```{r}

# This code prints the billboard dataset, which is built in with the tidyverse

billboard
```

```{r}

# Determine if the above data frame is tidy. If it isn't, write code to make it tidy

```

### Separate Columns

```{r}

# This code separates the month_year column that we created into seperate month and year columns.

crm_long <- crm_long |>
  separate(month_year, c("month", "year"), sep = "_")

crm_long
```

```{r}

# Report total purchases and average purchases per customer for every
# region-month in 2018. Display all 48 combinations, highest average first.
```

### Unite Columns

```{r}

# This code unites the month and year back together.

crm_long |>
  unite("month_year", c(month, year), sep = "_")
```

```{r}

# This code uses lubridate to help R recognize the dates in our dataset. We add a default day value of 1 to each row in the data to accomplish this.

crm_long <- crm_long |>
  mutate(day = 1) |> 
  unite("date", c(day, month, year), sep = "-") |> 
  mutate(date = dmy(date))

crm_long
```

```{r}

# Plot average monthly purchases per customer in each region.

crm_long |> 
  group_by(date, region) |> 
  summarize(
    total_transactions = sum(transactions),
    avg_transactions = mean(transactions),
    .groups = "drop"
  ) |> 
  ggplot(aes(x = date, y = avg_transactions, color = region)) +
  geom_line() +
  scale_x_date(date_breaks = "month", date_labels = "%b")
```

### Data Classes and Types

```{r}

# This creates an atomic vector. What storage type do all three values share?

vector_example <- c(1, 2, "three")

vector_example
```

```{r}

# Different columns can have different types: integer IDs and character states.

tibble(id = 1:3, state = "AZ")
```

```{r}

# This creates a one-column tibble; its values are still character.
# Converting the vector to a tibble does not recover its original numeric types.

as_tibble(vector_example)
```

### Your Turn!

Suppose you are building a ggplot visualization that shows the distributions of customer incomes for different zip codes. However, R is treating the values of zip code as a numeric value. Is this correct? If not, what data type might make more sense? Perform the appropriate coercion.

```{r}

# This code saves the zip codes.

zip_codes <- c(30301, 60614, 90210, 10001, 73301, 15213, 44101, 84101, 19104, 96815)

glimpse(zip_codes)
```

```{r}

# Use this space to coerce the zip codes into a more appropriate type.

```

Store identifiers such as `"02138"` as character from the outset. Once a leading zero is lost in numeric storage, `as.character()` cannot restore it.

```{r}

# Create a factor whose levels run from the most common region to the least.
# geom_bar() counts customers in each region automatically.

crm_data |> 
  mutate(region = fct_infreq(region)) |> 
  ggplot(aes(x = region)) +
  geom_bar()
```

```{r}

# Reverse the factor-level order used by the plot; the row order is unchanged.

crm_data |> 
  mutate(region = region |> fct_infreq() |> fct_rev()) |> 
  ggplot(aes(x = region)) +
  geom_bar()
```

### Live Coding Exercise

Patagonia wants a plot that tells a useful story about customer purchase activity over time using these historical records.

- Which measure and time period would be useful to the client?
- How should we interpret an increase in purchase activity?
- Which assumptions or limitations could change our interpretation?

```{r}

# Use this space to participate in the live coding exercise.
```
